CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review to improve reproducibility

### 2021-02-23 codecheck-paper.tex
few more abbrevs

\section*{Keywords}

@@

\clearpage

@@

\textbf{Step 3:} The codechecker runs the code, based on instructions
provided by the author. They check if some or all of the results from
the paper can be reproduced. If there are any problems running the
code, the codechecker asks the author for help, updates, or further
documentation.  The burden to provide reproducible material lies with
the author.  The codechecker then tries to run the code again.  This
process iterates until either the codechecker is successful, or the
codechecker concludes the workflow is not reproducible.  As part of
this process, the codechecker could work entirely locally, relying on
their own computing resources, or in the cloud, e.g.,~using the open
MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives,
some of which are more tailored to scientific publications while
others offer commercial options for, e.g.,~publishers
(cf. \cite{konkol_publishing_2020}).  A cloud-based infrastructure
allows for the codechecker and author to collaboratively improve the
code and enforces a complete definition of the computing environment;
but, unless secure infrastructure is provided, e.g.,~by the publisher,
this requires the code and data to be published openly online.  Note
that the task of the codechecker is to check the ``mechanics'' of the
workflow not to evaluate its correctness.  In the context of
mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish
between \emph{verification} and \emph{validation}; following their
definition, a CODECHECK ensures verification
%\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
of computational results, i.e.,~checking that code generates the
output it claims to create, but not a validation, i.e.,~checking that
the code implements the right algorithm to solve the specific research
problem.  Nevertheless, simply attempting to reproduce an output may
highlight a submission's shortcomings in meeting a journal's
requirements (cf.~\cite{christian_journal_2020}) and may effectively
increase transparency, thereby improving practices
(cf.~\cite{nosek_scientific_2012}) even if the check does not go into
every detail.

\textbf{Step 4:} The codechecker writes a certificate stating how the
code was run and includes a copy of outputs (e.g.,~data, figures, or
tables) that were independently generated.  The certificate may
include further recommendations on how to improve the material.  The
free text in the certificate can describe what was really checked,
because each workflow is unique.  Since no specific tool or platform
should be required, such that no authors are excluded, it would be
futile for the codechecker to use automation or fixed checklists.

\textbf{Step 5:} The certificate and auxiliary files created during
the check, e.g.,~a specification of a computing environment, data
subsets or helper scripts, and the original code and data get
deposited in an open archive unless restrictions (data size, license
or sensitivity) apply.  Currently, the codecheckers deposit the
material on Zenodo themselves, but the publisher may complete this
step after integrating CODECHECK into its review process.  A badge or
other visual aid may be added to the deposit and the paper and link to
the certificate.  Although a badge simplifies the CODECHECK into a
binary value and risks introducing confusion regarding the extent of
the check, especially across disciplines and methods, a badge can
provide recognition value and acknowledges the completed CODECHECK.
The badge and the actual check are incentives for undertaking the
effort needed to provide a reproducible workflow.

\textbf{Step 6:} The publisher can, depending on the timing, provide
the certificate to peer-reviewers or editors or publish it and link
between certificate, paper, and any repositories through
metadata. Currently, the codechecker creates these connections on
Zenodo. They appear as links with a relationship type on the Zenodo
landing page for a record in the right-hand sidebar, e.g., the
``related identifiers'' and ``alternate identifiers'' of certificate
\texttt{2020-025} \cite{cert-2020-025}.  The publisher also credits
the codechecker for their work by depositing the activity in scholarly
profiles, such as ORCID (see peer review contributions in ORCID
records,
\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}).
The publisher also ensures proper publication metadata, e.g.,~links
from the certificate repository to the published paper or the original
code repository.

@@

We have a curated list of 25 certificates (Table~\ref{tab:register}).
These CODECHECKs fall into three themes: (1) classic and current
papers from computational neuroscience, (2) COVID-19 modelling
preprints, and (3) GIScience.

@@

We use freely available infrastructure, GitHub and Zenodo, to run our
system.  The \texttt{codecheckers} \textbf{GitHub} organisation at
\url{https://github.com/codecheckers} contains projects for managing
the project website, the codecheckers community and its discussions,
code repositories, and the main register of CODECHECKs. Both the
project website \url{https://codecheck.org.uk/} and the register at
\url{https://codecheck.org.uk/register} are hosted as GitHub pages.
The register database is a single table in CSV format that connects
the certificate identifier with the repository associated with a
CODECHECK. Each of these repositories, which currently can be hosted
on GitHub or Open Science Framework, contains the CODECHECK
metadata file \texttt{codecheck.yml}. The register further contains a
column for the type of check, e.g.,~community, journal, or conference,
and the respective GitHub issue where communications and assignments
around a specific check are organised. No information is duplicated
between the register and the metadata files. The continuous
integration infrastructure of GitHub, GitHub Actions, is used to
automate the register for viewing (HTML) and for integration with
other services (CSV, JSON).  \textbf{Zenodo} is our preferred open
repository for storing certificates. It mints DOIs for deposits and
ensures long-term availability of all digital artefacts related to the
project. The CODECHECK community on Zenodo is available at
\url{https://zenodo.org/communities/codecheck/}. It holds
certificates, the regularly archived register
\cite{codecheck_register_jan2021}, and other material related to
CODECHECK.

@@

These tools and resources demonstrate that a CODECHECK process can be
managed without infrastructure costs on freely available platforms.
Automation of some aspects may improve turnaround time. Our main
resource requirements are the \textbf{humans} needed for managing the
project and processes and the codecheckers.  All contributions
currently rely on (partly grant-based) public funding and
volunteering.

@@

The journal \emph{ACM Transactions on Mathematical Software (TOMS)}
recently established a ``Replicated Computational Results'' (RCR)
review process \cite{heroux_editorial_2015}, where ``replicable'' is
the same as our use of ``reproducible''.  Fifteen RCR Reports have
been published so far (search on \url{https://search.crossref.org/}
with the term \texttt{"Replicated Computations Results (RCR) Report"}
on 2020-12-10).
% 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
and the process is being extended extended to the ACM journal
\emph{Transactions on Modeling and Computer Simulation (TOMACS)}.
%% (cf. \url{https://dl.acm.org/journal/tomacs/author-guidelines}).
The TOMS RCR follows CODECHECK principles 1-4, although our work was
independently developed of theirs.  The TOMS editorial
\cite{heroux_editorial_2015} shares similar concerns about selection
of reviewers, as we discussed above. Unlike existing CODECHECK
certificates, the RCR reports undergo editorial review.  Publication
of the RCR report recognises the efforts of the reproducing person,
while the potential for this motive to be a conflict of interest is
acknowledged.  TOMS also recognises reviewer activity in a partnership
with Publons (see
\url{https://authors.acm.org/author-services/publons}).

@@

Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline
and infrastructure for reproducible computational research with its
own tools and API. Their specific packaging format, provenance record,
and cryptographic \emph{Verifiable Result Identifier} would indeed
provide excellent reproducibility. However, the system is also complex
and since its creation in 2011 we are not aware of any publisher using
it; also, the system is not open source.  In comparison,
CODECHECK is less powerful but also much more flexible and less
dependent on specific tools or infrastructure. If data and code are
deposited properly, i.e.,~very unlikely to disappear, then the
certificate's DOI is practically close to the cryptographic
identifier.

@@

The question of training codecheckers is also relevant. We expect a
mentoring scheme within the CODECHECK community (experienced
codecheckers will provide on-the-job training or serve as fallback
advisors). Codecheckers may also be found by collaborating with
reproducible research initiatives such as
ReproHack (\url{https://reprohack.github.io/reprohack-hq/}),
ReproducibiliTea (\url{https://reproducibilitea.org/},
\cite{fitzgibbon_brewing_2020}), and
Repro4Everyone (\url{https://repro4everyone.org/},
\cite{auer_reproducibility_2020}).  The initial reaction of the
scientific community to these ideas shows that scholarly peer review
should continue on the path towards facilitating the sharing and
execution of computational workflows.

>>>>

\subsection*{Keywords}

@@

\subsection*{Abbreviations}

ACM: Association for Computing Machinery; 
ECRs: Early Career Researchers;
RCR: Replicated Computational Results;
TOMS: Transactions on Mathematical Software;
\clearpage

@@

\textbf{Step 3:} The codechecker runs the code, based on instructions
provided by the author. They check if some or all of the results from
the paper can be reproduced. If there are any problems running the
code, the codechecker asks the author for help, updates, or further
documentation.  The burden to provide reproducible material lies with
the author.  The codechecker then tries to run the code again.  This
process iterates until either the codechecker is successful, or the
codechecker concludes the workflow is not reproducible.  As part of
this process, the codechecker could work entirely locally, relying on
their own computing resources, or in the cloud, e.g.,~using the open
MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives,
some of which are more tailored to scientific publications while
others offer commercial options for, e.g.,~publishers
(cf. \cite{konkol_publishing_2020}).  A cloud-based infrastructure
allows for the codechecker and author to collaboratively improve the
code and enforces a complete definition of the computing environment;
but, unless secure infrastructure is provided, e.g.,~by the publisher,
this requires the code and data to be published openly online.  Note
that the task of the codechecker is to check only the ``mechanics'' of the
workflow.  In the context of
mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish
between \emph{verification} and \emph{validation}; following their
definition, a CODECHECK ensures verification
%\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
of computational results, i.e.,~checking that code generates the
output it claims to create, but not a validation, i.e.,~checking that
the code implements the right algorithm to solve the specific research
problem.  Nevertheless, simply attempting to reproduce an output may
highlight a submission's shortcomings in meeting a journal's
requirements (cf.~\cite{christian_journal_2020}) and may effectively
increase transparency, thereby improving practices
(cf.~\cite{nosek_scientific_2012}) even if the check does not go into
every detail.

\textbf{Step 4:} The codechecker writes a certificate stating how the
code was run and includes a copy of outputs (figures or tables) that
were independently generated.  The certificate may include
recommendations on how to improve the material.  The free text in the
certificate can describe exactly what was checked, because each
workflow is unique.  Since no specific tool or platform is required,
such that no authors are excluded, it is futile for the
codechecker to use automation or fixed checklists.

\textbf{Step 5:} The certificate and auxiliary files created during
the check, e.g.,~a specification of a computing environment, data
subsets or helper scripts, and the original code and data get
deposited in an open archive unless restrictions (data size, license
or sensitivity) apply.  Currently, codecheckers deposit the material
on Zenodo themselves, but a publisher may complete this step after
integrating CODECHECK into its review process.  A badge or other
visual aid may be added to the deposit and the paper and link to the
certificate.  Although a badge simplifies the CODECHECK into a binary
value and risks introducing confusion regarding the extent of the
check, a badge provides recognition value and acknowledges the
completed CODECHECK.  The badge and the actual check are incentives
for undertaking the effort needed to provide a reproducible workflow.

\textbf{Step 6:} The publisher can, depending on the timing, provide
the certificate to peer-reviewers or editors or publish it and link
between certificate, paper, and any repositories. Currently, the
codechecker creates these connections on Zenodo. They appear as links
with a relationship type on the Zenodo landing page for a certificate,
e.g., the ``related identifiers'' and ``alternate identifiers'' of
certificate \texttt{2020-025} \cite{cert-2020-025}.  The publisher
also credits the codechecker's work by depositing the activity
in scholarly profiles, such as ORCID (see peer review contributions in
ORCID records,
\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}).
The publisher also ensures proper publication metadata, e.g.,~links
from the certificate repository to the published paper or the original
code repository.

@@

To date we have created 25 certificates (Table~\ref{tab:register}) 
falling into three broad themes: (1) classic and current
papers from computational neuroscience, (2) COVID-19 modelling
preprints, and (3) GIScience.

@@

We use freely available infrastructure, GitHub and Zenodo, to run our
system.  The \texttt{codecheckers} \textbf{GitHub} organisation at
\url{https://github.com/codecheckers} contains projects for managing
the project website, the codecheckers community and its discussions,
code repositories, and the main register of CODECHECKs. Both the
project website \url{https://codecheck.org.uk/} and the register at
\url{https://codecheck.org.uk/register} are hosted as GitHub pages.
The register database is a single table in CSV format that connects
the certificate identifier with the repository associated with a
CODECHECK. Each of these repositories, which currently can be hosted
on GitHub or Open Science Framework, contains the CODECHECK
metadata file \texttt{codecheck.yml}. The register further contains a
column for the type of check, e.g.,~community, journal, or conference,
and the respective GitHub issue where communications and assignments
around a specific check are organised. No information is duplicated
between the register and the metadata files. The continuous
integration infrastructure of GitHub, GitHub Actions, is used to
automate generation of the register.  \textbf{Zenodo} is our preferred open
repository for storing certificates. It mints DOIs for deposits and
ensures long-term availability of all digital artefacts related to the
project. The CODECHECK community on Zenodo is available at
\url{https://zenodo.org/communities/codecheck/}. It holds
certificates, the regularly archived register
\cite{codecheck_register_jan2021}, and other material related to
CODECHECK.

@@

These tools and resources demonstrate that a CODECHECK process can be
managed on freely available platforms.
Automation of some aspects may improve turnaround time. Our main
resource requirements are the \textbf{humans} needed for managing the
project and processes and the codecheckers.  All contributions
currently rely on (partly grant-based) public funding and
volunteering.

@@

The journal \emph{ACM Transactions on Mathematical Software (TOMS)}
recently established a ``Replicated Computational Results'' (RCR)
review process \cite{heroux_editorial_2015}, where ``replicable'' is
the same as our use of ``reproducible''.  Fifteen RCR Reports have
been published so far (search on \url{https://search.crossref.org/}
with the term \texttt{"Replicated Computations Results (RCR) Report"}
on 2020-12-10).
% 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
and the process is being extended extended to the ACM journal
\emph{Transactions on Modeling and Computer Simulation}.
%% (cf. \url{https://dl.acm.org/journal/tomacs/author-guidelines}).
The TOMS RCR follows CODECHECK principles 1-4, although our work was
independently developed of theirs.  The TOMS editorial
\cite{heroux_editorial_2015} shares similar concerns about selection
of reviewers, as we discussed above. Unlike existing CODECHECK
certificates, the RCR reports undergo editorial review.  Publication
of the RCR report recognises the efforts of the reproducing person,
while the potential for this motive to be a conflict of interest is
acknowledged.  TOMS also recognises reviewer activity in a partnership
with Publons (see
\url{https://authors.acm.org/author-services/publons}).

@@

Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline
and infrastructure for reproducible computational research. Their
specific packaging format, provenance record, and cryptographic
\emph{Verifiable Result Identifier} would indeed provide excellent
reproducibility. However, the system is also complex and since its
creation in 2011 we are not aware of any publisher using it; also, the
system is not open source.  In comparison, CODECHECK is less powerful
but also much more flexible and less dependent on specific tools or
infrastructure. If data and code are deposited properly, i.e.,~very
unlikely to disappear, then the certificate's DOI is practically close
to the cryptographic identifier.

@@

The question of training codecheckers is also relevant. We expect a
mentoring scheme within the CODECHECK community (experienced
codecheckers will provide on-the-job training or serve as fallback
advisors). Codecheckers may also be found by collaborating with
reproducible research initiatives such as
ReproHack (\url{https://reprohack.github.io/reprohack-hq/}),
ReproducibiliTea (\url{https://reproducibilitea.org/},
\cite{fitzgibbon_brewing_2020}), and
Repro4Everyone (\url{https://repro4everyone.org/},
\cite{auer_reproducibility_2020}).  The initial reaction of 
researchers to these ideas shows that scholarly peer review
should continue on the path towards facilitating sharing and
execution of computational workflows.
